\documentclass[a4paper, 10pt]{article}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[colorlinks = true,linkcolor = blue,urlcolor  = blue,citecolor = blue,anchorcolor = blue]{hyperref}
\usepackage[usenames]{color}
 \urlstyle{same}
\usepackage{enumitem}

\usepackage{geometry}
\geometry{a4paper,left=1.2in,right=1.2in,top=1in,bottom=1in}

\usepackage{palatino}
\setlist[itemize]{leftmargin=1em}
\setlist[enumerate]{leftmargin=1em}

%\leftmargin=0.25in
%\oddsidemargin=0.25in
%\textwidth=6.0in
%\topmargin=0.25in
%\textheight=9.25in

\raggedright

\pagenumbering{arabic}

\def\bull{\vrule height 0.8ex width .7ex depth -.1ex }
% DEFINITIONS FOR RESUME

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}
}

\newcommand{\lineover}{
	\begin{changemargin}{-0.05in}{-0.05in}
		\vspace*{-8pt}
		\hrulefill \\
		\vspace*{-2pt}
	\end{changemargin}
}

\newcommand{\header}[1]{
	\begin{changemargin}{-0.5in}{-0.5in}
		\scshape{#1}\\
  	\lineover
	\end{changemargin}
}

\newcommand{\contact}[5]{
	\begin{changemargin}{-0.5in}{-0.5in}
		\begin{center}
			{\Large \scshape {#1}}\\ \smallskip
			{#2}\\ \smallskip
			{#3}\\ \smallskip
			{#4}\\ \smallskip
			{#5}\smallskip
		\end{center}
	\end{changemargin}
}

\newenvironment{body} {
	\vspace*{-16pt}
	\begin{changemargin}{-0.3in}{-0.5in}
  }	
	{\end{changemargin}
}	

\newcommand{\school}[4]{
	\textbf{#1} \hfill \emph{#2\\}
	#3\\
	#4\\
}

% END RESUME DEFINITIONS

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Name
\contact{\bf{Cong Ma}
\vspace{6pt}}{\emph{Address: } Jones 313, 5747 South Ellis Avenue,
Chicago, IL 60637}{\emph{Homepage:} \url{https://congma1028.github.io/}}{\emph{Email:} congm@uchicago.edu}



\header{\LARGE{Employment}}
\begin{body}
	\vspace{18pt}
	\textbf{Assistant professor} 
	 \hfill \emph{July~2021 - present} \\
	\begin{itemize}
		\item[-] Department of Statistics and the College, University of Chicago
	\end{itemize}
	
		\textbf{Postdoctoral scholar}\hfill \emph{June~2020 - June~2021} \\
    \begin{itemize}
      \item[-] Department of Statistics and Department of Electrical Engineering and Computer Sciences, University of California, Berkeley
      \item[-] Advisor: Professor \emph{Martin J.~Wainwright}
      %\item GPA: 3.976 / 4.0
    \end{itemize}
\end{body}


\bigskip
\header{\LARGE{Education}}
\begin{body}
	\vspace{18pt}

	\textbf{Ph.D.} in \textbf{Operations Research and Financial Engineering} \hfill \emph{May 2020}
    \begin{itemize}
      \item[-] Princeton University, advised by Professor \emph{Yuxin Chen} and Professor \emph{Jianqing Fan}
      %\item GPA: 3.976 / 4.0
    \end{itemize}
    \textbf{B.E.} in \textbf{Electrical Engineering} \hfill \emph{July 2015} \\
    \begin{itemize}
      \item[-] Tsinghua University
    \end{itemize}
    \textbf{Exchange student} \hfill \emph{Aug.~2013 - Dec.~2013} \\
    \begin{itemize}
      \item[-] School of Electrical and Computer Engineering, {Georgia Institute of Technology}
    \end{itemize}
\end{body}



\bigskip
\header{\LARGE{Research Interests}}
\begin{body}
\vspace{18pt}
Mathematics of data science, reinforcement learning, high-dimensional statistics, large-scale optimization, inference and uncertainty quantification, and their applications to neuroscience and signal processing.
\end{body}



\bigskip
\header{\LARGE{Preprints}}
\begin{body}
	\vspace{18pt}
	\begin{enumerate}[label={[{P}{{\arabic*}}]}]

\item \textbf{C.~Ma}, R.~Pathak, M.~J.~Wainwright, ``Optimally Tackling Covariate Shift in RKHS-based Nonparametric Regression'', 2022.

\item Y.~Yang, \textbf{C.~Ma}, {``$O(T^{-1})$ Convergence of Optimistic-Follow-the-Regularized-Leader in Two-Player Zero-Sum Markov Games,''} 2022. \\

	
\item Y.~Yang, \textbf{C.~Ma}, {``Optimal Tuning-Free Convex Relaxation for Noisy Matrix Completion,''} 2022. \\




	\end{enumerate}
\end{body}


\bigskip
\header{\LARGE{Journal articles}}
\begin{body}
\vspace{18pt}
\begin{enumerate}[label={[{J}{{\arabic*}}]}]


\item T.~Tong, \textbf{C.~Ma}, A.~Prater-Bennette, E.~Tripp, Y.~Chi, {``Scaling and Scalability: Provable Nonconvex Low-Rank Tensor
Estimation from Incomplete Measurements,''} accepted to \emph{Journal of Machine Learning Research}, 2022. \\

\item P.~Rashidinejad, B.~Zhu, \textbf{C.~Ma}, J.~Jiao, S.~Russel, {``Bridging Offline Reinforcement Learning and Imitation Learning: A Tale of Pessimism,''} accepted to \emph{IEEE Transactions on Information Theory}, 2022. 

\item \textbf{C.~Ma}, B.~Zhu, J.~Jiao, M.~J.~Wainwright, {``Minimax Off-Policy Evaluation for Multi-Armed Bandits,''} \emph{IEEE Transactions on Information Theory}, vol.~68, no.~8, pp.~5314 - 5339, Mar.~2022. \\

\item Y.~Chen, Y.~Chi, J.~Fan, \textbf{C.~Ma} (\textbf{alphabetical order}), {``Spectral Methods for Data Science: A Statistical Perspective,''}  \emph{Foundations and Trends in Machine Learning}, 2021. \\

\item Y.~Chen, J.~Fan, \textbf{C.~Ma}, Y.~Yan (\textbf{alphabetical order}), {``Bridging Convex and Nonconvex Optimization in Robust PCA: Noise, Outliers, and Missing Data,''}  \emph{Annals of Statistics}, vol.~49, no.~5, pp.~2948-2971, 2021. \\

\item Y.~Chen, \textbf{C.~Ma}, H.~V.~Poor, Y.~Chen, {``Learning Mixtures of Low-Rank Models,''} \emph{IEEE Transactions on Information Theory}, vol.~67, no.~7, pp.~4613-4636, July 2021. \\






\item T.~Tong, \textbf{C.~Ma}, Y.~Chi, {``Low-Rank Matrix Recovery with Scaled Subgradient Methods: Fast and Robust Convergence Without the Condition Number,''}  \emph{IEEE Transactions on Signal Processing}, vol.~69, no.~3, pp.~2396-2409, 2021. \\







\item J.~Fan, \textbf{C.~Ma}, Y~.Zhong (\textbf{alphabetical order}), {``A Selective Overview of Deep Learning,''} \emph{Statistical Science}, vol.~36, no.~2, pp.~264-290, May 2021 (\textbf{invited overview article}).\\

\item T.~Tong, \textbf{C.~Ma}, Y.~Chi, {``Accelerating Ill-Conditioned Low-Rank Matrix Estimation via Scaled Gradient Descent,''} \emph{Journal of Machine Learning Reseasrch}, vol.~22, no.~150, pp.~1-63, May 2021. \\

\item Y.~Li, \textbf{C.~Ma}, Y.~Chen, Y.~Chi, {``Nonconvex Matrix Factorization from Rank-One Measurements,''} \emph{IEEE Transactions on Information Theory}, vol.~67, no.~3, pp.~1928-1950, March 2021.

\item \textbf{C.~Ma}, Y.~Li, Y.~Chi, {``Beyond Procrustes: Balancing-Free Gradient Descent for Asymmetric Low-Rank Matrix Sensing,''} \emph{IEEE Transactions on Signal Processing}, vol.~69, pp.~867-877, Jan.~2021.


	
\item	\textbf{C.~Ma}, J.~Lu, H.~Liu, {{``Inter-Subject Analysis: A Partial Gaussian Graphical Model Approach,''}}  \emph{Journal of the American Statistical Association}, vol.~116, no.~534, pp.~746-755, 2021. \\

\item Y.~Chen, Y.~Chi, J.~Fan, \textbf{C.~Ma}, Y.~Yan (\textbf{alphabetical order}), {``Noisy Matrix Completion: Understanding Statistical Guarantees for Convex Relaxation via Nonconvex Optimization,''} \emph{SIAM Journal on Optimization}, vol.~30, no.~4, pp.~3098–3121, Oct.~2020. \\

\item \textbf{C.~Ma}, K.~Wang, Y.~Chi, Y.~Chen, {{``Implicit Regularization in Nonconvex Statistical Estimation: Gradient Descent Converges Linearly for Phase Retrieval, Matrix Completion, and Blind Deconvolution,''}}  \emph{Foundations of Computational Mathematics}, vol.~20, no.~3, pp.~451-632, June
2020. \\


\item Y.~Chen, Y.~Chi, J.~Fan, \textbf{C.~Ma} (\textbf{alphabetical order}), {``Gradient Descent with Random Initialization: Fast Global Convergence for Nonconvex Phase Retrieval,''} \emph{Mathematical Programming,} vol.~176, no.~1-2, pp.~5-37, July 2019. \\


\item Y.~Chen, J.~Fan, \textbf{C.~Ma}, Y.~Yan (\textbf{alphabetical order}), {``Inference and Uncertainty Quantification for Noisy Matrix Completion,''} \emph{Proceedings of the National Academy of Sciences~(PNAS)} vol.~116, no.~46, pp.~22931-22937, Nov.~2019 (direct submission). \\




	
\item 
	Y.~Chen, J.~Fan, \textbf{C.~Ma}, K.~Wang (\textbf{alphabetical order}), {{``Spectral Method and Regularized MLE Are Both Optimal for Top-$K$ Ranking,''}} \emph{Annals of Statistics,} vol.~47, no.~4, pp.~2204-2235, Aug.~2019. \\







	
	\item J.~Zhang, J.~Tang, \textbf{C.~Ma}, H.~Tong, Y.~Jing, J.~Li, W.~Luyten, M-F.~Moens, {{``Fast and Flexible Top-$k$ Similarity Search on Large Networks,''}} \emph{ACM Transactions on Information Systems}, vol.~36, no.~2,  pp.~13:1-13:30, Sept.~2017.\\
	
\end{enumerate}
\end{body}

\bigskip
\header{\LARGE{Conference papers}}
\begin{body}
	\vspace{18pt}
	\begin{enumerate}[label={[{C}{{\arabic*}}]}]
	
	\item R. Pathak, C. Ma, M. J. Wainwright, {``A New Similarity Measure for Covariate Shift with Applications to Nonparametric Regression''}, \emph{International Conference on Machine Learning (ICML)}, 2022 (long presentation).
	
	\item G.~Li, \textbf{C.~Ma}, N.~Srebro (\textbf{alphabetical order}), {``Pessimism for Offline Linear Contextual Bandits using $\ell_p$ Confidence Sets, ''} \emph{Conference on Neural Information Processing Systems (Neurips)}, 2022.
	\item P.~Rashidinejad, B.~Zhu, \textbf{C.~Ma}, J.~Jiao, S.~Russel, {``Bridging Offline Reinforcement Learning and Imitation Learning: A Tale of Pessimism,''} \emph{Conference on Neural Information Processing Systems (Neurips)}, 2021. \\
	
	\item \textbf{C.~Ma}, Y.~Li, Y.~Chi, {``Beyond Procrustes: Balancing-Free Gradient Descent for Asymmetric Low-Rank Matrix Sensing,''} \emph{Asilomar Conference on Signals, Systems and Computers}, Nov.~2019. \\
	\item 
	Y.~Li, \textbf{C.~Ma}, Y.~Chen, Y.~Chi, {{``Nonconvex Matrix Factorization from Rank-One Measurements,''}} \emph{International Conference on Artificial Intelligence and Statistics (AISTATS)}, Apr.~2019. \\
	
	\item 
	\textbf{C.~Ma}, K.~Wang, Y.~Chi, Y.~Chen, {{``Implicit Regularization in Nonconvex Statistical Estimation: Gradient Descent Converges Linearly for Phase Retrieval and Matrix Completion,''}} \emph{International Conference on Machine Learning (ICML)}, July 2018. \\
	
	\item J.~Zhang, J.~Tang, \textbf{C.~Ma}, H.~Tong, Y.~Jing, J.~Li, {{``Panther: Fast Top-$k$ Similarity Search on Large Networks,''}} \emph{ACM SIGKDD Conference on Knowledge Discovery and Data Mining (KDD)}, Aug.~2015.\\
	
	
	\end{enumerate}
\end{body}


\bigskip
\header{\LARGE{Tutorials}}
\begin{body}

\vspace{18pt}
\begin{enumerate}[label={[{Tut}{{\arabic*}}]}]
\item  {``Nonconvex Optimization for High-Dimensional Signal Estimation: Spectral and Iterative Methods,''} European Signal Processing Conference (EUSIPCO) 2020, together with Y.~Chen and Y.~Chi.
\end{enumerate}

\end{body}


\bigskip
\header{\LARGE{Invited talks}}
\begin{body}
	\vspace{18pt}
	\begin{enumerate}[label={[{T}{{\arabic*}}]}]
	\item ``Bridging Offline Reinforcement Learning and Imitation Learning: A Tale of Pessimism,'' \emph{Statistics Seminar, University of Bristol,} Nov.~2022.
	
	\item ``Bridging Offline Reinforcement Learning and Imitation Learning: A Tale of Pessimism,'' \emph{SIAM Conference on Mathematics of Data Science,} Sept.~2022.
	
	\item ``Minimax Off-Policy Evaluation for Multi-Armed Bandits,'' \emph{Statistics and Data Science Seminar, UIC, } Dec.~2021. 
	
	\item ``Bridging Offline Reinforcement Learning and Imitation Learning: A Tale of Pessimism,'' \emph{Joint CS and TTIC Machine Learning Seminar}, Oct.~2021.
	
	\item ``Bridging Offline Reinforcement Learning and Imitation Learning: A Tale of Pessimism,'' \emph{Informs Annual Meeting}, Oct.~2021.
	
	\item ``Minimax Off-Policy Evaluation for Multi-Armed Bandits,'' \emph{FODSI Seminar, UC Berkeley}, Apr.~2021. 
	\item ``Bridging Convex and Nonconvex Optimization in Noisy Matrix Completion: Stability and Uncertainty Quantification,'' \emph{BLISS Seminar, UC Berkeley}, Nov.~2020.
	\item ``Bridging Convex and Nonconvex Optimization in Noisy Matrix Completion: Stability and Uncertainty Quantification,'' \emph{Young Data Science Researcher Seminar, ETH}, June~2020.
	\item ``Statistics Meets Nonconvex Optimization: Computational Efficiency and Uncertainty Quantification,'' \emph{ISyE Seminar, School of Industrial \& Systems Engineering, Georgia Tech, }Feb.~2020.
	\item ``Statistics Meets Nonconvex Optimization: Computational Efficiency and Uncertainty Quantification,'' \emph{Department Seminar, Department of IEOR, Columbia University, }Feb.~2020.
	\item ``Statistics Meets Nonconvex Optimization: Computational Efficiency and Uncertainty Quantification,'' \emph{Department Seminar, Department of Statistics, University of Pennsylvania, }Feb.~2020.
	\item ``Statistics Meets Nonconvex Optimization: Computational Efficiency and Uncertainty Quantification,'' \emph{Department of Statistics, UC Irvine, }Feb.~2020.
	\item ``Statistics Meets Nonconvex Optimization: Computational Efficiency and Uncertainty Quantification,'' \emph{Department Seminar, Department Seminar, Department of Statistics, UC Davis, }Feb.~2020.
	\item ``Statistics Meets Nonconvex Optimization: Computational Efficiency and Uncertainty Quantification,'' \emph{Department of Statistics, Rutgers University, }Jan.~2020.
	\item ``Statistics Meets Nonconvex Optimization: Computational Efficiency and Uncertainty Quantification'', \emph{Statistics Colloquium, Department of Statistics, University of Chicago,} Jan.~2020.
	\item ``Statistics Meets Nonconvex Optimization: Computational Efficiency and Uncertainty quantification,'' \emph{Department Seminar, Department of Data Sciences and Operations, USC}, Jan.~2020.
	\item ``Nonconvex Optimization Meets Statistics: Towards Rigorous Computational and Inferential Guarantees,'' \emph{School of Industrial Engineering, Purdue University}, Dec.~2019.
	\item ``Bridging convex and nonconvex optimization in noisy matrix completion: Stability and uncertainty quantification,'' \emph{Statistics Seminar, Department of Mathematics, University of Maryland, College Park}, Dec.~2019.
	\item ``Understanding statistical properties by bridging convex and nonconvex optimization,'' \emph{AMS Fall Eastern Sectional Meeting, Binghamton University}, Oct.~2019.
	\item ``Noisy matrix completion: understanding statistical guarantees for convex relaxation via nonconvex optimization,'' \emph{Annual Young Researchers Workshop}, \emph{ORIE, Cornell University}, Oct.~2019.
	\item ``Noisy matrix completion: understanding statistical guarantees for convex relaxation via nonconvex optimization,'' \emph{Mathematics, Information and Computation (MIC) Seminar}, \emph{Center for Data Science and Courant Institute, NYU}, Apr.~2019.
		\item ``Nonconvex Matrix Completion Without Regularization,'' \emph{Asilomar Conference on Signals, Systems, and Computers, Pacific Grove}, Oct.~2018.
		\item ``Spectral Method and Regularized MLE are Both Optimal for Top-$K$ Ranking,'' \emph{Energy and Information Systems Seminar, CMU}, July 2018.
		\item ``Gradient Descent with Random Initialization: Fast Global Convergence for Nonconvex Phase Retrieval,'' \emph{ICML Workshop on Modern Trends in Nonconvex Optimization for Machine Learning, Stockholm}, July 2018.
		\item ``Implicit Regularization in Nonconvex Statistical Estimation: Gradient Descent Converges Linearly for Phase Retrieval, Matrix Completion'', \emph{International Conference on Machine Learning (ICML), Stockholm}, July 2018.
		\item ``Implicit Regularization in Nonconvex Statistical Estimation,'' \emph{International Symposium on Mathematical Programming, Bordeaux}, July 2018.
		\item ``Implicit Regularization in Nonconvex Statistical Estimation,'' \emph{Department of Electrical Engineering, Tsinghua University}, Jan.~2018.
			
		\item ``Spectral Method and Regularized MLE are Both Optimal for Top-$K$ Ranking,'' \emph{International Conference on Data Science, Shanghai}, Dec.~2017.
		\item ``Inter-Subject Analysis: Inferring Sparse Interactions with Dense Intra-Graphs,'' \emph{ICSA Applied Statistics Symposium, Chicago}, June 2017.
		
		\end{enumerate}
\end{body}
%
\bigskip
\header{\LARGE{Professional Service}}
\begin{body}
	\vspace{18pt}
	\begin{enumerate}
	\item Co-organizer of \emph{Statistics Colloquium}, Department of Statistics, University of Chicago \hfill Sept. 2018 -- present
	\item Co-organizer of \emph{Wilks statistics seminar}, ORFE, Princeton University \hfill July 2018 -- May 2019
	\item Co-organizer of \emph{$6^{\mathsf{th}}$ Princeton Day of Statistics}, ORFE, Princeton University
%	\item Co-organizer of reading group on mathematical data science, Princeton University \hfill Sept.~2018 -- May~2020
	\item Member of Technical Program Committee for \emph{52nd Annual Conference on Information Sciences and Systems} (\emph{CISS 2018})
	\item Reviewer for the following journals: \emph{Annals of Statistics}, \emph{Statistical Science}, \emph{Bernoulli Journal}, \emph{Operations Research}, \emph{Proceedings of the IEEE}, \emph{Journal of Machine Learning Research}, \emph{IEEE Transactions on Signal Processing}, \emph{IEEE Transactions on Information Theory}, \emph{Journal of Business \& Economic Statistics}, \emph{Transactions on Knowledge and Data Engineering}, \emph{SIAM Journal on Mathematics of Data Science}
	\item Reviewer for the following conferences: \emph{ACM Symposium on Theory of Computing}, \emph{Conference on Neural Information Processing Systems} (\emph{NeurIPS}), \emph{International Conference on Artificial Intelligence and Statistics}, \emph{IEEE International Symposium on Information Theory}, \emph{Annual Conference on Information Sciences and Systems}, \emph{International Conference on Acoustics, Speech, and Signal Processing} (\emph{ICASSP})

	\end{enumerate}
\end{body}

%
\bigskip
\header{\LARGE{Teaching}}
\begin{body}
	\vspace{18pt}
	\begin{enumerate}
	\item Topics in Learning Under Distribution Shifts, Winter 2023
	\item Introduction to Probabilistic Models, Winter 2022, 2023
	\item Topics in Mathematical Data Science: Spectral Methods and Nonconvex Optimization, Fall 2021 
%	\item \emph{Statistical Foundations of Data Science} by Professor J.~Fan, Spring 2019, Head Teaching Assistant
%	\item \emph{Mathematics of High-Dimensional Data} by Professor Y.~Chen, Fall 2018, Head Teaching Assistant
%	\item \emph{Large-Scale Optimization for Data Science} by Professor Y.~Chen, Spring 2018, Head Teaching Assistant
%	\item \emph{Statistical Learning and Nonparametric Estimation} by Professor Han Liu, Spring 2017, Teaching Assistant
%	\item \emph{Analysis of Big Data} by Professor Han Liu, Spring 2017, Teaching Assistant
%	\item \emph{Probability and Stochastic Systems} by Professor Mykhaylo Shkolnikov, Fall 2016, Teaching Assistant
	\end{enumerate}
\end{body}


%\bigskip
%\header{\LARGE{Internship Experience}}
%\begin{body}
%\vspace{18pt}
%\emph{AI Labs, Hudson River Trading, New York, NY} \hfill \emph{June~2019 - Aug.~2019} \\
%
%\end{body}

\bigskip

\header{\LARGE{Awards and Honors}}

\begin{body}
    \vspace{18pt}
    \begin{enumerate}
    \item  Hannan Graduate Student Travel Award, Institute of Mathematical Statistics \hfill{} \emph{2020} \\
    \item {School of Engineering and Applied Science Award for Excellence}, Princeton University \hfill{} \emph{2019}\\
    {\em This award is given to SEAS advanced graduate students who have performed at the highest level as scholars and~researchers.} \\
    \smallskip
    \item {AI Labs Fellowship}, Hudson River Trading  \hfill{} \emph{2019}\\
    \smallskip
   \item {Best Poster Award}, Princeton Day of Optimization  \hfill{} \emph{2018}\\
%    \smallskip
%     \item{Student Travel Award}, $35^{\mathsf{th}}$ International Conference on Machine Learning (ICML)  \hfill{} \emph{2018}\\
%    \smallskip
%     \item{SEAS Travel Grant}, Princeton University  \hfill{} \emph{2018}\\
%     \smallskip
%     \item{SEAS Travel Grant}, Princeton University  \hfill{} \emph{2017}\\
     \smallskip
     \item {ICSA Student Paper Award}, International Chinese Statistical Association  \hfill{} \emph{2017}\\
     \smallskip
     \item  {First Year Engineering Fellowship}, Princeton University \hfill{} \emph{2015}\\
    \smallskip
    \item {Outstanding Academic Performance Scholarship}, Tsinghua University  \hfill{} \emph{2014}\\
    \smallskip
    \item {National Scholarship}, Tsinghua University (Highest honor, 2 out of 118 in department)\hfill{} \emph{2013}\\
    \smallskip
    \item {CNPC Scholarship}, Tsinghua University (Highest honor, only recipient in  department)\hfill{} \emph{2012}
    \end{enumerate}
\end{body}


\end{document}
