\documentclass{article}
\usepackage{url}

\usepackage{cite,enumitem,amsmath, amsfonts, amssymb}
\usepackage{epsfig,subfigure}
\usepackage{comment}
\usepackage{array}


\newcommand\dueDate{11:59pm on Oct. 21st}

\input assignment_utils
\usepackage{listings}


\begin{document}
\createHomework{1}
%\createHomeworkSolutions{1}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}{Concentration of gaussian random variable}{20}

\newpart{5}
Let $X$ be a standard normal random variable. Prove that 
\[
\mathbb{P} ( |X| \leq t) \leq 2 \exp(- t^2 / 2). 
\]

\newpart{5}
Let $X_1, X_2, \ldots, X_n$ be $n$ i.i.d.~standard normal random variables. Prove that with probability at least $1- O(n^{-10})$, one has
\[
\max_{i}  |X_i|  \leq 5 \sqrt{\log n}.
\]

\newpart{10}
Let $\bm{x} \in \mathbb{R}^{n}$ be a random vector where each coordinate is an independent standard normal random variable. Using the the conclusion above, one can show that $\|\bm{x}\|_{2} \lesssim \sqrt{n \log n}$ with high probability. However, this falls short in two aspects. First, the upper bound on $\|x\|_2$ is not tight. Second, it doesn't provide a high-probability lower bound of $\|x\|_2$. In this part, prove that for all $t \in (0,1)$, one has
\[
\mathbb{P}( | \| \bm{x} \|_{2}^2 - n | \geq nt) \leq 2 \exp(-nt^2 / 8).
\]
(Hint: Laplace transform method with control of moment generating function)
\end{problem}

\begin{problem}{Norm of Gaussian random matrices}{40}

Recall that in class, we have used the bound $\| \bm{E} \| \lesssim \sqrt{n}$ where $\bm{E} \in \mathbb{R}^{n \times n}$ is composed of i.i.d.~standard normal random variables. 

\newpart{10} Use matrix Bernstein's inequality to show that with high probability $\|\bm{E}\| \lesssim \sqrt{n \log n}$. (Hint: truncation) \\

As before, the bound proved in part (a) is off by a $\sqrt{\log n}$ factor. In the following, we will prove the tighter version. Recall the definition of $\|\bm{E}\|$:
\[
\|\bm{E}\| = \sup_{\| \bm{v}\|_2 =1} \|\bm{E} \bm{v} \|_2
\]
Hence it suffices to show that with high probability $\sup_{\| \bm{v}\|_2 =1} \|\bm{E} \bm{v} \|_2 \lesssim \sqrt{n}$.


\newpart{5} 
Let's first focus on a fixed vector $\| \bm{v}\|_2 =1 $. Prove that for any fixed $\bm{v} \in \mathbb{R}^{n}$, one has 
\[
\mathbb{P}( \|\bm{E} \bm{v} \|_2 \geq 10\sqrt{n}) \leq 2 \exp(-100n).
\] 

It is tempting to apply ``union bound'' and ``obtain''
\[
\mathbb{P}( \sup_{\| \bm{v}\|_2 =1} \|\bm{E} \bm{v} \|_2 \geq 10\sqrt{n}) \leq \sum_{\| \bm{v}\|_2 =1} \mathbb{P}( \|\bm{E} \bm{v} \|_2 \geq 10\sqrt{n}).
\]

However this argument is TOTALLY wrong as one cannot apply union bound to a uncountable set. Therefore, to properly apply union bound, one needs to restrict attention to a finite subset of the unit sphere in $\mathbb{R}^{n}$ that can approximate the unit sphere well. \\

\newpart{10} Let $\mathcal{N}_{\varepsilon}$ be a subset of $\{\bm{v} \in \mathbb{R}^{n}: \| \bm{v}\|_2 =1\}$ such that for any point $\bm{v}$ in $\{\bm{v} \in \mathbb{R}^{n}: \| \bm{v}\|_2 =1\}$, one can find an element $\bm{u} \in \mathcal{N}_{\varepsilon}$ such that $\|\bm{u} - \bm{v}\|_{2} \leq \varepsilon$. In particular, set $\mathcal{N}_{\varepsilon}$ be such a set with smallest cardinality. Prove that 
\[
|\mathcal{N}_{\varepsilon}| \leq (1 + \tfrac{2}{\varepsilon})^{n},
\]
where $|\mathcal{N}_{\varepsilon}|$ denotes the cardinality of $\mathcal{N}_{\varepsilon}$.

\newpart{5}
Fix some $\varepsilon \in (0,1)$. Prove that for any matrix $\bm{A} \in \mathbb{R}^{n \times n}$
\[
\| \bm{A} \| \leq \frac{1}{ 1 - \varepsilon } \cdot \max_{\bm{v} \in \mathcal{N}_{\varepsilon} }\|\bm{A} \bm{v}\|_{2}.
\] 
This shows the usefulness of $\mathcal{N}_{\varepsilon}$ in terms of approximating $\|\bm{A}\|$.

\newpart{10} Combine the previous steps to show that with high probability $\|\bm{E}\| \lesssim \sqrt{n}$.
\end{problem}

\begin{problem}{Matrix concentration in matrix completion}{10}
	Consider the matrix completion problem introduced in class where $\bm{M}^{\star} = \bm{U}^\star \bm{\Sigma}^{\star} \bm{V}^{\star \top} \in \mathbb{R}^{n \times n}$ is a rank-$r$ matrix. Let $\mu$ be its incoherence parameter. Prove that with high probability 
	\[
	\|\bm{M} - \bm{M}^{\star}\| \lesssim \sqrt{\frac{\mu r \log n}{n p}} \|\bm{M}^\star\|
	\]
	as long as $n p \geq C \mu r \log n$ for some sufficiently large constant $C > 0$.
\end{problem}

\begin{problem}{Community detection experiments}{20}
Consider the SBM model discussed in class. Fix the number $n$ of nodes in a graph to be $100$. Set $p=\frac{1+ \varepsilon}{2}$ and $q=\frac{1-\varepsilon}{2}$ for some quantity $\varepsilon \in [0,1/2]$. Generate a random graph and then use the spectral method to cluster the nodes. Please plot the mis-clustering rate vs. the probability gap $\varepsilon$. At the minimum, you should take 50 different values of $\varepsilon$ (with linear spacing) in $[0, 1/2]$. For each value of $\varepsilon$, you need to run the experiment with at least 200 Monte-Carlo trials to calculate the average mis-clustering rate across trials. 
\end{problem}

\begin{problem}{Matrix completion experiments}{20}
Consider the SBM model discussed in class. Fix the number $n$ of nodes in a graph to be $100$. Set $p=\frac{1+ \varepsilon}{2}$ and $q=\frac{1-\varepsilon}{2}$ for some quantity $\varepsilon \in [0,1/2]$. Generate a random graph and then use the spectral method to cluster the nodes. Please plot the mis-clustering rate vs. the probability gap $\varepsilon$. At the minimum, you should take 50 different values of $\varepsilon$ (with linear spacing) in $[0, 1/2]$. For each value of $\varepsilon$, you need to run the experiment with at least 200 Monte-Carlo trials to calculate the average mis-clustering rate across trials. 
\end{problem}





\end{document}
